\Large \bf FlexAlloc: Dynamic Memory Partitioning\\ between the Corevisor and Hostvisor in SeKVM

### 2023-12-24 paper.tex
added evaluation edits

Another key shortcoming of our design is that it does not fully implement
dynamic memory partitioning between the corevisor and hostvisor. As mentioned in
the introduction, another area in which corevisor memory scales with the number
of VMs is in maintaining \textit{struct el2\_vm\_info} objects that record the
metadata associated with each guest VM. Currently, we still statically allocate
a fixed-size array of \textit{struct el2\_vm\_info} objects in corevisor
memory ~\cite{hypsec_host.h}. This essentially preserves the hard cap on the
number of guest VMs that can be run. An easy fix would be to keep this array
as a statically-allocated array, while increasing the number of elements. Since each
\textit{struct el2\_vm\_info} object is only 768 bytes, this would not waste a
significant amount of memory even if fewer than sixteen VMs are created.

@@

We also do not fully evaluate our implementation. For one, we do not stress-test our implementation by running a large number
of VMs. To do this, we would have to modify the macros used to define the current
sixteen-VM limit. It is entirely conceivable that the hostvisor might not be able
to allocate 1M contiguous regions of memory past a certain point if memory
pressure and external fragmentation are high. In this case, the hostvisor would
not be able to allocate the memory regions needed for the corevisor to create
the stage-2 page table for a new VM. We could perhaps delay this limit by breaking
the VM stage-2 page tables down into even smaller regions of memory, but this would
likely lead to more complicated logic when traversing the page table. In addition, we do not verify whether fragmenting the stage-2 page table results
in any degradation in VM performance. Ideally, we would run a memory-intensive
benchmark on our guest VMs to verify that our iterators work properly under a
high memory load. Finally, we do not quantify the change in corevisor memory usage.
We assume that by not allocating the memory for VM stage-2 page tables at bootup,
corevisor memory will automatically be reduced by an amount matching the
size of this static allocation. This should be verified using actual data. For
instance, we could periodically iterate over every page and see how many pages
are owned by the corevisor before and after VM creation.

@@

A possible area of future work is exploring other allocators designed for
allocating large chunks of contiguous memory. We fragment the VM stage-2 page tables
into eight distinct 1M pages, because \textit{alloc\_pages} does not allow us
to allocate contiguous regions larger than 4M. It might be possile to avoid this
by looking at other Linux memory allocators explicitly designed for the purpose
of large contiguous allocations.

To enable dynamic allocation of EL2 memory in general beyond just the scope of
a VM's stage-2 page tables, we also designed but did not implement
a dynamic page manager (DPM). DPM is a system that exposes an interface for other ``services'' within EL2 to
utilize dynamic allocation for their memory regions, while hiding the implementation of actually receiving memory from
and returning memory to the host. This would be another interesting area of future work.

>>>>

Another key shortcoming of our design is that it does not fully implement
dynamic memory partitioning between the corevisor and hostvisor outside of stage-2 page tables for VMs. As mentioned in
the introduction, another area in which corevisor memory scales with the number
of VMs is in maintaining \textit{struct el2\_vm\_info} objects that record the
metadata associated with each guest VM. Currently, we still statically allocate
a fixed-size array of \textit{struct el2\_vm\_info} objects in corevisor
memory ~\cite{hypsec_host.h}. This essentially preserves the hard cap on the
number of guest VMs that can be run. An easy fix would be to keep this array
as a statically-allocated array, while increasing the number of elements. Since each
\textit{struct el2\_vm\_info} object is only 768 bytes, this would not waste a
significant amount of memory even if fewer than sixteen VMs are created. Regarding the
\textit{shared\_data\_region}, since each VM requires a \textit{struct kvm}, the size of
this region would either need to be dynamically allocated or the static size would
need to be increases in order to lift the 16 VM cap present in SeKVM. However, since the static region
size has been decreased by 128 MB within FlexAlloc and the size of \textit{struct kvm}'s is relatively
small, inflating this region to accomodate 32+ VMs would still result in substantial memory savings
thanks to FlexAlloc's changes. In the Future Work section, we propose a potential solution to
the issue of bringing dynamic allocation to EL2 metadata structures that are not page aligned
which we call DPM.

@@

We also do not fully evaluate our implementation. For one, we do not stress-test our implementation by running a large number
of VMs. To do this, we would have to modify the macros used to define the current
sixteen-VM limit. It is entirely conceivable that the hostvisor might not be able
to allocate 1M contiguous regions of memory past a certain point if memory
pressure and external fragmentation are high. In this case, the hostvisor would
not be able to allocate the memory regions needed for the corevisor to create
the stage-2 page table for a new VM. We could perhaps delay this limit by breaking
the VM stage-2 page tables down into even smaller regions of memory. In addition, we do not verify whether fragmenting the stage-2 page table results
in any degradation in VM performance. Given the less-than-ideal test environment
we have at our disposal, we did not investigate performance metrics related to execution time
for our implementation and only considered memory footprint metrics. Ideally, we would run a memory-intensive
benchmark on our guest VMs to verify that our iterators work properly under a
high memory load across multiple VMs.

@@

A possible area of future work is exploring other allocators designed for
allocating large chunks of contiguous memory. We fragment the VM stage-2 page tables
into eight distinct 1M pages, because \textit{alloc\_pages} does not allow us
to allocate contiguous regions larger than 4M. We also wanted to be reasonably certain
that the size of our regions would not impeed the host's ability to launch VMs. It might be possile to avoid this
by looking at other Linux memory allocators explicitly designed for the purpose
of large contiguous allocations and adding such an allocator to the host.

\subsection{Dynamic Page Manager}
To enable dynamic allocation of EL2 memory in general beyond just the scope of
a VM's stage-2 page tables, we also designed but did not fully implement
a dynamic page manager (DPM). DPM is a system that exposes an interface for other ``services'' within EL2 to
utilize dynamic allocation for their memory regions, while hiding the implementation of actually receiving memory from
and returning memory to the host. This would be another interesting area of future work.
